---
title: "Hmk_05"
author: "Elaine Nunan"
format: gfm
editor: visual
---

# Question 1: filtering

We were tasked with making a plot of x = air time and y = distance. However, the distance had to meet extra criteria:

-   originate from LaGuardia airport ("LGA")

-   departed on the 16th of the month

-   have a flight distance of less than 2000

    ```{r, message= FALSE}
    library(nycflights13)
    library(tidyverse)

    flight_filtered <- filter(flights, flights$origin == "LGA", flights$day == 16, flights$distance < 2000)

    ggplot(data = flight_filtered) +
      geom_point(aes(x=air_time, y=distance))
    ```

I loaded the `nycflights13` using the `library()` function as well as the `tidyverse`. Then I filtered the `flights` data frame by the `origin` being set to LaGuardia, the `day` of the month being set to the 16th, and the `distance` being set to less than 2000 miles. I assigned this filtered data frame to the object "flight_filtered". I created a scatter plot using the filtered data frame and set the aesthetics to x = `air_time` and y = `distance`. The plot worked! However, there was a message that 75 points were omitted from the graph due to missing values.

# Question 2: dealing with NAs

Now we are tasked with making a data frame that has values for every row in the `arr_time` and `dep_time`.

```{r}
schedule_times <- filter(flights, arr_time >= 0, dep_time >= 0)

times_plot <- filter(flights, origin == "LGA", day == 16, distance < 2000, air_time >= 0)

ggplot(data = times_plot) +
  geom_point(aes(x=air_time, y=distance))
```

To make this data frame, I filtered the `flight` data frame according to `arr_time` and `dep_time` being greater than or equal to 0, which should catch all of the flights as long as they have a value inputted. As for being able to prevent the warning message about ggplot not being able to graph all the points due to missing values, I tried making a new data frame (I called it "times_plot") that includes `air_time` in the filter. This way the filter will take out any rows that have NA in the `air_time` column.

# Question 3: adding columns

Now, we are tasked with creating a new with creating a data frame of average flight speeds using `air_time` and `distance` then creating a histogram or a density plot of the new `flight_speeds` data.

```{r}
b <- mutate(flights, speed = distance / air_time)

ggplot(data = b) +
  geom_histogram(aes(x=speed))

```

I created the new column, called `speed` in the `flight` data frame by using the `mutate()` function and passing it the arguments `distance` divided by `air_time`. I assigned this new data frame to the object "b" then created a histogram and assigned the x aesthetic `speed`.
