---
title: "Homework 10: Decisions and loops"
author: "Elaine Nunan"
format: gfm
editor: visual
---

# Decisions

Write a function that accepts the current time as a parameter and prints "Good morning", "Good afternoon", or "Good evening" depending on the time. It is fine for the time to be in numeric format (e.g., `2317` for 11:17 pm).

```{r, message=FALSE}
library(tidyverse)

watch <- function(time) {
if(time >= 000 & time <= 1100) {
  print ("Good morning!")
} else if( time >= 1101& time <= 1600) {
  print("Good afternoon!")
} else if( time >= 1601 & time <= 2359) {
  print ("Good evening")
} else {
   print("You aren't using military time, you noob")
}
}

time <- 2400
watch(time)
```

I created a function called `watch` (as in the timepiece) to state the appropriate greeting based on inputted time (in military time). If a time that isn't in the correct format or outside the normal 24 hour limit then the code will print a (gentle) reminder to use military time.

# Loops

-   Write a `for` loops that calculates the mean of each column of mtcars

```{r}
  cars.means <- rep(NA, length(mtcars))
for(i in seq_along(mtcars)) {
  cars.means[i] <- mean(mtcars[ ,i])
} 
```

I initialized the results vector `cars.means` so that it will have the appropriate number of columns. Then I created a `for` lop that for every ith element of `cars.means` to input the mean of every ith column of `mtcars`. This loop works because all the columns in the `mtcars` data set is numeric.

-   Write a function (using a for loop) that calculates the mean of all numeric columns of *any* data frame. This function should be able to accept data frames with non-numeric columns.

```{r}

ifnum_mean <- function(df) {
  calc.mean <- rep(NA, length(df))
  for(i in seq_along(df)) {
    if(is.numeric(df [ ,i])) {
      calc.mean[i] <- mean(df [ ,i])
    } else if (!is.numeric(df [ ,i])) {
      calc.mean[i] <- NA 
    } 
  }
  calc.mean
}

iris <- as.data.frame(iris)
ifnum_mean(iris)
```

My function `ifnum_mean` assesses the columns of the data frame to see if the columns are numeric or non-numeric. If they are numeric then it calculates the mean then if it isn't it inputs NA.

## Why not loops

In R, we generally encourage people to use vectorized functions instead of `for` loops. According to [your textbook](https://r4ds.had.co.nz/iteration.html), what is better about vectorized functions?

`For` loops aren't the easiest thing to read or use whereas vectorized functions are usually named for the process they do. `For` loops also require as extra step of initializing the results vector whereas vectorized functions have that built-in.
